\section{Live-Testing}

Despite the poor results from \autoref{chap:backtesting}, a strategy will be executed live on a demo account at ByBit for further work.
This will serve, among other things, to compare the live-executed strategy with the backtest results and to test the robustness of the trading engine in live operation.

\subsection{Broker Setup}

To connect to ByBit the \verb|ByBit-Broker-Connector| and \verb|ByBit-Remote-DataSource| components have been used.

The \verb|ByBit-Remote-DataSource| is able to:

\begin{enumerate}
    \item Request historical market data by the ByBit \texttt{Get Kline} API Endpoint \cite{get-kline}.
    \item Stream market data by the ByBit \texttt{Kline} websocket endpoint \cite{stream-kline}.
\end{enumerate}

The \verb|ByBit-Broker-Connector| is able to execute following actions via the ByBit API:

\begin{enumerate}
    \item Request the current wallet balance as well as the current margin \cite{wallet-balance}.
    \item Placing Orders in the market, which includes open, take-profit, and stop-loss orders. Each Order is supported as limit and market order \cite{place-order}.
    \item Request all open positions for one or multiple pairs \cite{position-info}. In the current paper this only includes positions on \ethusdc.
    \item Request closed trades in a specified time period \cite{closed-pnl}.
\end{enumerate}

The connection to ByBit is established by an API key and an API secret which can be created in the ByBit web interface.
The distinction between a demo account and a real money account is regulated by ByBit via the base URL, which is different in each case.
To use the demo depot, for REST endpoints, the base url \texttt{https://api-testnet.bybit.com} has to be used.
For the real money account, the base url \texttt{https://api.bybit.com} has to be used \cite{integration}.
To stream the market data, no distinction is required because the base URL \texttt{wss://stream.bybit.com/v5/public/linear} \cite{ws-connect} for streaming USDC market data is a public endpoint, which can be used without API Key.
In the trading engine, the base URLs, API key, and API secret are set by environment variables to achieve an independent implementation, which is configured by the user.

\subsection{Live-Test Results}

Using the trading engine configuration described above, a live execution of the triple exponential moving average strategy was performed from \liveStartDataStartDate to \liveStartDataEndDate.
To keep the environment as similar as possible to the local backtest environment, the balance of the demo account was set to 5000 USDC at startup.